---
title: "Vector exercises"
output: html_notebook
---

# For loop exercises

Here is your template for a `for` loop.  This is a loop to repeat something `n` times (where `n` is a variable).  Let's say we want to repeat something 10 times.  We first set `n` to 10, then do our `for` loop, like this:

```{r}
n <- 10
for (i in 1:n) {
  # The thing you want to repeat
}
```

You may remember you can print a number (or anything else) with the `print` function, like this:

```{r}
print(5)
```

Use a `for` loop to print out all the numbers from 1 through 6, one number on each line.  Fill in the `...` parts, and the loop body, after the comment.

```{r}
for ( ... in ... ) {
  # The body of the loop
  ...
}
```

Use a `for` loop to print out all the numbers from 3 through 13, one number on each line.  Fill in the `...` parts, and the loop body, after the comment.


```{r}
for ( ... in ... ) {
  # The body of the loop
  ...
}
```

One way of keeping a running total, is to set a variable to be itself, plus another number.  For example, to add the values 3, 5 and 7 together, we could do:

```{r}
my_var <- 0
my_var <- my_var + 3
my_var <- my_var + 5
my_var <- my_var + 7
my_var
```

Use a `for` loop to add all the numbers 5 through 20.   Show the result at the end:

```{r}
total <- 0
for ( ... in ...) {
  # The body of the loop
  ...
}
# Show the result
total
```

You should get the same result as:

```{r}
sum(5:20)
```

Here is the code to make a vector of 1000 random numbers between 0 and 1.

```{r}
# Make the vector, store it in the "r" variable
r <- runif(1000)
# Show the first six values
head(r)
```

Show a histogram of 1000 random numbers between 0 and 1.

```{r}
hist(...)
```

Here is the sum of 50 random numbers:

```{r}
sum(runif(50))
```

In the chunk below, you start with a vector of 1000 zeros.  Use a for loop to replace each zero in the vector with the sum of a fresh set of 50 random numbers.   Show a histogram of the values

```{r}
# A vector of 1000 zeros
sum_of_50 <- numeric(1000)
for ( ... in ...) {
  sum_of_50[...] <- ...
}
hist(sum_of_50)
```

If you were surprised by the shape of the result, well done, you have just demonstrated the [Central Limit Theorem](https://en.wikipedia.org/wiki/Central_limit_theorem).  When you add random numbers, whatever their distribution, the distribution of the result tends to get closer to the normal distribution.

Here is how you get the maximum value from a vector:

```{r}
values <- c(3, 1, 4, 1, 5, 9, 2, 6)
max(values)
```

Take 1000 vectors of 50 random numbers, and record the maximum for each
vector.  Do a histogram of the result:

```{r}
# A vector of 1000 zeros
max_of_50 <- numeric(1000)
for ( ... in ...) {
  ... <- ...
}
hist(max_of_50)
```

Notice that finding the maximum does not involve adding, so the central limit theorem does not apply.

What distribution do you think you will get from the *median* of 50 random numbers?  You can use the `median` function to get the median value from a vector:

```{r}
# Make a vector of 1000 zeros
...
for ( ... ) {
  # Calculate median and insert into vector
   ... <- ...
}
hist(...)
```

As you see in the chunk below, you start with the vector `v <- c(3, 1, 4, 1, 5)`.  Your job is to use a `for` loop to change `v` so the first element (`v[1]`) is 3 * 1, the second element (`v[2]`) is 3 * 2, the third is 4 * 3, the fourth is 1 * 4 and the fifth is 5 * 5.

```{r}
v <- c(3, 1, 4, 1, 5)
for ( ... in ... ) {
  # The body of the loop
  v[...] <- ...
}
# Show the result
v
```

The `for` loop above is just for practice.  You would usually do the same thing by multiplying vectors, like this:

```{r}
v <- c(3, 1, 4, 1, 5)
# Elementwise multiplication
v * 1:5
```

Make sure you understand why this is the same.
